Single-cell virus sequencing of influenza infections that trigger innate immunity

### 2018-12-27 paper/paper.tex
Viral variation feels colloquial, tried a slightly different wording
just fixing a minor typo

But although viral variation is clearly important, it still only partially explains heterogeneity among influenza-infected cells.
We find substantial breadth in viral transcriptional burden and occasional IFN induction even among cells infected with unmutated virions.
Additionally, none of the viral defects in IFN+ cells induce IFN deterministically---and the most immunostimulatory defect (absence of NS) also occurs in multiple IFN- cells in our single-cell dataset.
Of course, our single-cell virus sequencing only calls mutations present in a cell at relatively high frequency, and so could overlook low-frequency mutations that contribute to IFN induction.
However, we suspect that stochasticity or pre-existing cellular states explain most remaining heterogeneity, since IFN induction is heterogeneous even among cells treated with synthetic innate-immune ligands~\citep{shalek2013single, shalek2014single, wimmers2018single, bhushal2017cell} or other viruses~\citep{oneal2018west}.

Perhaps the most intriguing question is how the heterogeneity that we have described ultimately affects the macroscopic outcome of infection.
Natural human influenza infections are initiated by just a handful of virions~\citep{mccrone2018stochastic, xue2018reconciling, varble2014influenza} that then undergo exponential growth, and early IFN responses are amplified by paracrine signaling~\citep{stetson2006type,honda2006type}.
It is therefor eplausible that early heterogeneity could have a large effect on downstream events.
Extending our approaches to more complex systems could shed further light on how viral variation interacts with cell-to-cell heterogeneity to shape the race between virus and immune system.

>>>>

But although variation within viral populations is clearly important, it still only partially explains heterogeneity among influenza-infected cells.
We find substantial breadth in viral transcriptional burden and occasional IFN induction even among cells infected with unmutated virions.
Additionally, none of the viral defects in IFN+ cells induce IFN deterministically---and the most immunostimulatory defect (absence of NS) also occurs in multiple IFN- cells in our single-cell dataset.
Of course, our single-cell virus sequencing only calls mutations present in a cell at relatively high frequency, and so could overlook low-frequency mutations that contribute to IFN induction.
However, we suspect that stochasticity or pre-existing cellular states explain most remaining heterogeneity, since IFN induction is heterogeneous even among cells treated with synthetic innate-immune ligands~\citep{shalek2013single, shalek2014single, wimmers2018single, bhushal2017cell} or other viruses~\citep{oneal2018west}.

Perhaps the most intriguing question is how the heterogeneity that we have described ultimately affects the macroscopic outcome of infection.
Natural human influenza infections are initiated by just a handful of virions~\citep{mccrone2018stochastic, xue2018reconciling, varble2014influenza} that then undergo exponential growth, and early IFN responses are amplified by paracrine signaling~\citep{stetson2006type,honda2006type}.
It is therefore plausible that early heterogeneity could have a large effect on downstream events.
Extending our approaches to more complex systems could shed further light on how viral variation interacts with cell-to-cell heterogeneity to shape the race between virus and immune system.
